---
title: "Cournot Nash Equilibrium"
subtitle: "Industrial Organization and Competition Policy"
author: "Marly Tatiana Celis"
date: "December 7th, 2023"
output:  slidy_presentation 
# output:  beamer_presentation
# pdf_document:
#     fig_width: 7
#     fig_height: 6
#     fig_caption: true
# #  theme: "default"
# #     keep_tex: true
# # classoption: aspectratio=169    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(ggplot2)
library(tidyverse)
#tinytex::install_tinytex(force =TRUE)
```


## Cournot Market Structure

This is a simple code for representing the Cournot equilibrium for a duopoly when market demand is $Q(P)=200-P$.
Each firm’s cost function is $C(q_i )=20q_i$, where $i=1,2$. 

- In this case the Cournot model corresponds to two firms with symmetric costs.

- Each firm's goal is to chose the level of output that maximizes profits, given the output of the other firm.

- Firm i's payoffs are profits $\pi_i(q_i,q_j)=(p-c)q_i$


## Create data frame with output values for each firm

The first step consist on creating a data frame with two columns. On column for the output for firm 1 and another column with the output for firm 2.

Next, we create some value with the parameters of the demand and costs functions. 
```{r cournot, echo = TRUE, eval=TRUE}
rm(list=ls())

# Create output q and define the range of values
cournotdata <- data.frame(output_firm1 = seq(0, 180, by = 10),
                 output_firm2 = seq(0, 180, by = 10))

# Create values with parameters of the functions 
a <- 200
c <- 20
b <- 1


```

## Nash equilibrium

At the Nash equilibrium, recall that each firm must behave optimally assuming that its rival b ehaves optimally. That is, firm $i$ maximizes profits , believing that firm $j$ maximizes its profits. Another way of saying this is that each firm calculates its best response or reply to the
expected best-reply behavior of the other firm

$q_1=\frac{a-c-bq_2}{2b}$ 

$q_2=\frac{a-c-bq_1}{2b}$ 

## Compute best responses

```{r cournot data, echo=TRUE}

# Compute best responses
cournotdata <- cournotdata %>% 
  mutate( bestresp_Firm1 = (a-c-b*output_firm2)*((2*b)^-1),
          bestresp_Firm2 = (a-c-b*output_firm1)*((2*b)^-1)
          )

ggplot() +
  geom_line(data = cournotdata, aes(x = output_firm1, y = bestresp_Firm2), color = "blue", size = 1.5) +
  geom_line(data = cournotdata, aes(x = bestresp_Firm1, y = output_firm2), color = "red", size = 1.5) +
  labs(title = "Best Responses", x = "Output Firm 1", y = "Output Firm 2") +
  scale_x_continuous(breaks = seq(0, max(cournotdata$output_firm1, cournotdata$bestresp_Firm1), by = 20)) +
  scale_y_continuous(breaks = seq(0, max(cournotdata$output_firm1, cournotdata$bestresp_Firm1), by = 20)) +
  theme_minimal()


  
  
```


## Compute equilibrium


Solving simultaneaously for $q_1$ and $q_2$ we get 

$q_1 =\frac{a-c}{3b}$ 

$q_2 =\frac{a-c}{3b}$ 

Compute best responses

```{r cournot nash, echo = TRUE}

# Compute best responses

nashoutput_firm1= (a-c)*((3*b)^-1)
nashoutput_firm2= (a-c)*((3*b)^-1)

ggplot() +
  geom_line(data = cournotdata, aes(x = output_firm1, y = bestresp_Firm2), color = "blue", size = 1.5) +
  geom_line(data = cournotdata, aes(x = bestresp_Firm1, y = output_firm2), color = "red", size = 1.5) +
  labs(title = "Best Responses", x = "Output Firm 1", y = "Output Firm 2") +
  geom_point(aes(x = nashoutput_firm1, y = nashoutput_firm2), color = "green", size = 3) +  # Add the single point
  scale_x_continuous(breaks = seq(0, max(cournotdata$output_firm1, cournotdata$bestresp_Firm1), by = 20)) +
  scale_y_continuous(breaks = seq(0, max(cournotdata$output_firm1, cournotdata$bestresp_Firm1), by = 20)) +
  theme_minimal()


  
  
```

