---
title: "Assignment 1 - ECB3IO Problem 2(a) welfare and monopoloist"
author: ""
subtitle: "Industrial Organization and Competition Policy"
date: "`r Sys.Date()`"
output:  slidy_presentation 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(ggplot2)
library(tidyverse)
```

## Profit Maximizing Monopolist

Assignment 1 problem 2 requested a graphical representation in RStudio\footnote{Or other software} of the concepts of welfare and deadweight loss in the context of a monopoly.

This is a sample code to plot monopoly demand, marginal cost, marginal revenue, and the area of deadweight loss. 

Please read pages 33-36 of the book "Industrial Organization: A Strategic Approach" by Church and Ware (2000) for a thourghouh description of the concepts and formulas employed.

## Load necessary libraries

```{r setting, echo = TRUE}

library(tidyverse)

rm(list=ls())

# Set up data for a linear inverse demand curve
quantity <- seq(0, 10, by = 1)
a <- 10
b <- 0.5
inverse_demand <- 10 - 0.5 * quantity  # Linear inverse demand equation: P = 10 - 0.5Q
marginalcost <- 6
pricemonopoly <- (a+marginalcost)/2
quantitymonopoly <- (a-marginalcost)/(2*b)
quantitysocial <- (a-marginalcost)/(b)

# Create a data frame
df_inverse_demand <- data.frame(quantity, inverse_demand)

```

## Plot the linear inverse demand curve

```{r InverseDemand,  echo = TRUE}


# Plot the linear inverse demand curve
p_inverse_demand <- ggplot(df_inverse_demand, aes(x = quantity, y = inverse_demand)) +
  geom_line(color = "blue", size = 1, linetype = "solid") +
  # Add labels
  labs(title = "Linear Inverse Demand Curve",
       x = "Quantity",
       y = "Price") +
  # Customize the theme
  theme_minimal()
# Print the plot
print(p_inverse_demand)


```

## Example of inverse demand plus marginal revenue
```{r Monopoly, echo = TRUE}
# Calculate marginal revenue
slope_demand <- -0.5  # slope of the inverse demand curve
marginal_revenue <- 10 - (2*0.5)* quantity


# Add the marginal revenue variable to the data frame
df_inverse_demand$marginal_revenue <- marginal_revenue



```

## Plot the linear inverse demand curve with marginal revenue

```{r AllVariables, echo = TRUE}
# Plot the linear inverse demand curve with marginal revenue
p_inverse_demand <- ggplot(df_inverse_demand, aes(x = quantity)) +
  geom_line(aes(y = inverse_demand), color = "blue", size = 1, linetype = "solid", ) +
  geom_line(aes(y = marginal_revenue), color = "deepskyblue", size = 1, linetype = "dashed", ) +
  geom_line(aes(y = marginalcost), color = "black", size = 0.5, linetype = "solid", ) +
  geom_line(aes(y = pricemonopoly), color = "deeppink2", size = 0.5, linetype = "solid",) +
  geom_vline(xintercept = quantitymonopoly, linetype = "dashed",) +
  geom_vline(xintercept = quantitysocial, linetype = "dashed",) +
  theme_minimal()

print(p_inverse_demand)

``` 

## Add Deadweight loss


```{r Deadweightloss,  echo = TRUE}


# Add the Deadweight loss
p_inverse_demand <- p_inverse_demand +
  geom_polygon(data = data.frame(x = c(quantitymonopoly, quantitymonopoly, quantitysocial),
                                 y = c(pricemonopoly, marginalcost, marginalcost)),
               aes(x = x, y = y), fill = "gray", alpha = 0.5) +
  geom_text(x=(quantitymonopoly+1.5), y=(pricemonopoly-1.2), label="DWL") 

print(p_inverse_demand)

```

## Add labesl and text
```{r Additional,  echo = TRUE }
p_inverse_demand <- p_inverse_demand + ggtitle("Plot of Profit Maximizing Monopolist \n Deadweight loss") +
  xlab("Quantity") + ylab("Price") +
  geom_text(x=(quantitymonopoly-3), y=(pricemonopoly+0.2), label="Monopoly Price", color="deeppink2") +
  geom_text(x=(quantitymonopoly-3), y=(marginalcost+0.2), label="Marginal cost", color = "black") +
  geom_text(x=(quantitymonopoly), y=(pricemonopoly+1), label="Inverse Demand", color = "blue") + 
  geom_text(x=(quantitymonopoly+3), y=(pricemonopoly-3.5), label="Marginal Revenue", color = "deepskyblue")

# Print the plot
print(p_inverse_demand)

``` 


